---
title: "a6_task3_gabrielle_barsotti"
author: "Gabrielle Barsotti"
date: "12/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor) 
library(here)
library(broom)
library(kableExtra) 
library(stargazer) 
library(GGally)
library(ggbeeswarm)
library(effsize)
```

## Overview

In this document, a comparison was made between the lengths of female lizards of two species, Cnemidophorus uniparens, Desert Grassland lizard (CNUN) and Cnemidophorus tessalatus, Colorado Checkered lizard (CNTE). The data was explored using ggpairs analysis, histograms and GG-plots, and analysis of mean and standard deviation of the two species. Figure 1 provides a visual comparison and Table 1 provides the summary statistics. As a result of the data exploration, it was determined that it would be appropriate to run a two-sided t-test to understand the relationship between lizard species and total body length. Then Cohen's d effect size was calculated to determine the difference between mean lengths for the two species.

```{r, include = FALSE}

#Step 1: Read in data and clean the names

lizards <- read_csv(here("data", "lizards.csv")) %>%  
  clean_names()
```

```{r, include = FALSE}

#Step 2. Making a subset

lizards_subset <- lizards %>% # Created a data set for just species, sex length
  select(spp, sex, total_length) %>% 
  filter(sex =="F") %>% # This filters out ONLY Female
  filter(spp %in% c("CNUN","CNTE")) %>%
  filter(!total_length == "NA")
```

```{r, include = FALSE}

#Step 3. Exploratory GG Plot

 lizards_subset %>% 
  ggpairs(aes(color = spp))

# Data comparisons:
# Coral = Desert Grassland lizard and turquoise = Colorado Checkered lizard
# They have Different means and both species length seem normally distributed but with a possibly of being negatively skewed
# Desert may have more intra-population variance
# The sample size is really small for both, but especially small for the Desert Grassland lizard with has less than 30 samples
```

```{r, echo = FALSE}

#Step 4. Visualizations

# Visualize lizard length by species and create a beeswarm plot, then plot the data with mean and SE shown on top 

ggplot(data = lizards_subset, aes(x = spp, y = total_length)) + #Created ggplot using sex and length
  geom_beeswarm(aes(color = spp), show.legend = FALSE) + #Created beeswarm graph within the ggplot function and take out the legend
  stat_summary(fun=mean,   # Inserted summary statistics into the graph
               geom = "point",
               shape = 20,
               size = 4,
               color = "black",
               fill = "black") +
  geom_boxplot(fill = NA, width = 0.5, outlier.color = NA) +
  theme_minimal() +
  labs(x = "Species", # Labeled the graph with x and y axis and title
       y = "Lizard Length (mm)",
       title = "Female Lizard Length by Species")

```

Figure 1. Histogram shows the total counts of lizard length measurements by species - Desert Grassland lizard (CNUN) and Cnemidophorus tessalatus, the Colorado Checkered lizard (CNTE).

Table 1. Distribution of lizard length by species - Desert Grassland lizard (CNUN) and the Cnemidophorus tessalatus, the Colorado Checkered lizard (CNTE). 

```{r, include=FALSE}

#Step 5. Summary Statistics Table

lizards_summary <- lizards_subset %>% # Created a separate table to summarize the stats
  filter(spp == "CNTE" | spp == "CNUN") %>% 
  group_by(spp) %>% # Group by sex
  summarize(
    mean_length = round(mean(total_length),2),
    sd_length = round(sd(total_length),2),
    sample_size = n()) # Include the sample size in the summary statistics table
```

```{r, echo = FALSE}
 kable(lizards_summary, 
       col.names = c("Species", "Mean Length (mm)", "Standard Deviation", "Sample Size")) %>%
   kable_styling(full_width = FALSE) 
```


```{r, include = FALSE}

#Step 6. Further exploration

# Pull vectors of length for the two species

CNTE <- lizards_subset %>% # Created a new dataset for CNTE species
  filter(spp == "CNTE") %>%
  pull(total_length) # Pulled the length

CNUN <- lizards_subset %>% # Create a new dataset for CNUN species
  filter(spp == "CNUN") %>%
  pull(total_length) # Pulled the length

# Plot histograms & qq-plots to check assumptions
hist(CNTE) # Slightly skewed
qqnorm(CNTE) # Fairly linear
hist(CNUN) # Slightly skewed, less of a spread
qqnorm(CNUN) # Seems pretty linear
```


```{r, include=FALSE}

#Step 7. Running a t-test

# Exploring the data: Do not include this in the final report
lizards_ttest <- t.test(CNTE, CNUN) # Run a two sample t-test for the two species

lizards_ttest # This calls one format of the T-test (the one we are used to seeing in lecture)

# Get the tidy model results to call outputs for in-line referencing : 
lizards_ttest_tidy <- tidy(lizards_ttest)

lizards_ttest_tidy # This calls the tidy version of the T-test

#The t-test results show: Estimate = Difference between means, Estimate 1 = mean for male, Estimate 2 = mean for female, Statistic = t-value, Parameter = Degrees of Freedom 
```

```{r, include=FALSE}

#Step 8. Finding the effect size using Cohen's d

# Getting the effect size
lizards_effsize <- cohen.d(CNTE, CNUN)

lizards_effsize
```

## Summary 
On average, female Colorado Checkered lizards (CNTE) have a longer body length than female Desert Grassland lizards at the Jornada Basin LTER. (`r lizards_summary$mean_length[1]` ± `r lizards_summary$sd_length[1]` and `r lizards_summary$mean_length[2]` ± `r lizards_summary$sd_length[2]`, respectively; mean ± 1 standard deviation).

While the absolute difference in means is `r (lizards_summary$mean_length[1]) - (lizards_summary$mean_length[2])` mm (a `r round((lizards_summary$mean_length[1] - lizards_summary$mean_length[2])/(lizards_summary$mean_length[2] + lizards_summary$mean_length[1])*100, 2)`% difference in relation to the CNUN species), the difference in means is significant (Welch's two-sample t-test: t(`r round(lizards_ttest_tidy$parameter,2)`) = `r round(lizards_ttest_tidy$statistic,2)`, p > 0.008), and the effect size is large (Cohen's *d* = `r round(lizards_effsize$estimate,2)`).  (Figure 1 and Table 1).


Citation:

Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative.  [Link](https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e)

Metadata: [Link](https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-jrn.210007001.36)
